A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds

### 2013-06-11 paper/paper.tex
combined figures and tables, tiny bit of writing

\begin{figure}
	\centering
		\includegraphics[width=0.23\linewidth]{figures/hops.pdf}
    \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$.}
    	\label{fig:hop}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[width=0.30\linewidth]{figures/jump.pdf}
    \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
    \label{fig:jump}
\end{figure}

Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric. Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:jump_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively. We let $d(v,w)$ denote the \textbf{distance between vertices $v$ and $w$} obtained by minimizing over all paths containing edges as well as hops and jumps. Similarly, we let $diam(M)$ denote the \textbf{diameter of $M$} defined in terms of the distance function $d$. Let $\mathcal{P}$ denote the set of all the paths containing edges, hops or jumps.

@@

A small Python program was created to examine each manifold in the census and compute its diameter. See Figure \ref{fig:global_statistics} for a plot of the number of manifolds by topological type and diameter. As claimed, every manifold with diameter greater than $J$ is a 3-sphere. The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:global_statistics_edges}. Note that there are $3$-manifolds homeomor

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/global_statistics.png}
    \end{center}
    \caption{Distribution of diameters.}
    \label{fig:global_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/split_statistics.png}
    \end{center}
    \caption{Distribution of diameters for each type.}
    \label{fig:type_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    \includegraphics[width=0.6\linewidth]{figures/edge_diameters.png}
    \end{center}
    \caption{Distribution of diameters when only edges are considered, without hops or jumps.}
    \label{fig:global_statistics_edges}
\end{figure}

\begin{figure}
\centering
\begin{tabular} {| r | r | r | r | r | r | r |}
\hline
Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
\hline
\hline
$1$&  &  &  &   &3    & 3    \\
$H$&1 &  &  &10 &173  & 184  \\
$2$&  &  &1 &7  &401  & 409  \\
$J$&  &1 &1 &5  &2438 & 2445 \\
$3$&  &  &  &   &1060 & 1060 \\
$2H$&  &  &  &   &582  & 582  \\
$4$&  &  &  &   &93   & 93   \\
$2J$&  &  &  &   &11   & 11   \\
\hline
Total&1 &1 &2 &22   &4761   &4787  \\
\hline
\end{tabular}
\caption{The positively curved combinatorial 3-manifolds by diameter and topological type}
\label{fig:manifold_counts}
\end{figure}

\begin{figure}
\centering
\begin{tabular} {| r | r | r | r | r | r | r |}	
\hline
Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
\hline
\hline
$1$ &  &  &  &   &3    & 3    \\
$2$ & 1&  &1 &17 &574  & 593  \\
$3$ &  &1 &1 &5  &3498 & 3505 \\
$4$ &  &  &  &   &675  & 675   \\
$5$ &  &  &  &   &11   & 11   \\
\hline
Total&1 &1 &2 &22   &4761   &4787  \\
\hline
\end{tabular}
\caption{The positively curved combinatorial 3-manifolds by edge-diameter and topological type}
\label{fig:manifold_counts_edges}
\end{figure}

\section{Algorithm Used}

TO DO: BRIEF DISCUSSION OF ALGORITHM USED

>>>>

\begin{figure}
	\centering
		\includegraphics[width=0.28\linewidth]{figures/hops.pdf}
    \caption{A {\em hop} from vertex $v_1$ to vertex $v_2$}
    	\label{fig:hop}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[width=0.36\linewidth]{figures/jump.pdf}
    \caption{A {\em jump} from vertex $v_1$ to vertex $v_2$}
    \label{fig:jump}
\end{figure}

Just as for edges in an edge-path, the length of each hop and jump will be its length as a PL-path in the standard PL metric. Some Euclidean geometry tells us these lengths are \begin{equation} \label{eqn:hop_length} H = \frac{2}{3}\sqrt{2} \end{equation} and \begin{equation} \label{eqn:jump_length} J = \frac{1}{2}\sqrt{2} + \sqrt{3} \end{equation} respectively. We let $d(v,w)$ denote the \textbf{distance between vertices $v$ and $w$} obtained by minimizing over all paths containing edges as well as hops and jumps. Similarly, we let $diam(M)$ denote the \textbf{diameter of $M$} defined in terms of the distance function $d$. Let $\mathcal{P}$ denote the set of all the paths containing edges, hops or jumps.

@@

A small Python program was created to examine each manifold in the census and compute its diameter. See Figure \ref{fig:type_statistics} for a plot of the number of manifolds by topological type and diameter. As claimed, every manifold with diameter greater than $J$ is a 3-sphere. The corresponding data using the coarser-grained edge-diameter can be found in Figure \ref{fig:type_statistics_edges}. Note the existence of $3$-manifolds with edge-diameter greater than half the maximum allowed edge-diameter and which are {\em not} homeomorphic to the 3-sphere. In particular, there are positively curved combinatorial 3-manifolds homeomorphic to $\mathbb{RP}^3$, $L(3,1)$ and $L(4,1)$ which have edge-diameter three.

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{10pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.75\linewidth]{figures/split_statistics.png}}
    }
    \begin{tabular} {| r | r | r | r | r | r | r |}
		\hline
		Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$&  &  &  &   &3    & 3    \\
		$H$&1 &  &  &10 &173  & 184  \\
		$2$&  &  &1 &7  &401  & 409  \\
		$J$&  &1 &1 &5  &2438 & 2445 \\
		$3$&  &  &  &   &1060 & 1060 \\
		$2H$&  &  &  &   &582  & 582  \\
		$4$&  &  &  &   &93   & 93   \\
		$2J$&  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below)}
    \label{fig:type_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{10pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.75\linewidth]{figures/edge_diameters.png}}
    }
	\begin{tabular} {| r | r | r | r | r | r | r |}	
		\hline
		Edge-Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$ &  &  &  &   &3    & 3    \\
		$2$ & 1&  &1 &17 &574  & 593  \\
		$3$ &  &1 &1 &5  &3498 & 3505 \\
		$4$ &  &  &  &   &675  & 675   \\
		$5$ &  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below)}
	\label{fig:type_statistics_edges}
\end{figure}
